\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\title{DIP-002: Decision Integrity Protocol Specification\\
Canonical Payload and Verification Rules}

\author{Pavan Dev Singh Charak}

\date{2026}

\begin{document}

\maketitle

\begin{abstract}

This document specifies the canonical payload structure,
signing rules, and verification procedures used by the
Decision Integrity Protocol (DIP).

\end{abstract}

\section{Protocol Overview}

The Decision Integrity Protocol defines a deterministic
process for generating and verifying decision evidence.

The protocol includes the following stages:

\begin{enumerate}
\item Decision Execution
\item Payload Construction
\item Payload Canonicalization
\item Payload Signing
\item Ledger Append
\item Verification
\end{enumerate}

\section{Decision Payload Schema}

The canonical decision payload has the following structure:

\begin{verbatim}
{
  "protocol_version": "1.0",
  "decision_id": "string",
  "timestamp": "ISO-8601",
  "actor": "string",
  "action": "string",
  "inputs": {...},
  "result": "string",
  "signature": "ed25519"
}
\end{verbatim}

\section{Canonicalization}

Before signing, the decision payload MUST be canonicalized.

Canonicalization ensures that identical decision records
produce identical serialized payload representations.

\section{Signature Scheme}

Implementations SHOULD use the Ed25519 signature scheme.

The signing process:

\begin{enumerate}
\item Serialize canonical payload
\item Compute payload hash
\item Generate Ed25519 signature
\end{enumerate}

\section{Ledger Append}

Signed payloads MUST be appended to an append-only
decision ledger.

The ledger ensures chronological ordering and
immutability of decision evidence.

\section{Verification Procedure}

Verification of a decision record involves:

\begin{enumerate}
\item Parse decision payload
\item Reconstruct canonical payload
\item Verify cryptographic signature
\item Confirm ledger inclusion
\end{enumerate}

\section{Protocol Versioning}

Each payload MUST include a protocol version field
to support forward compatibility.

\section{Conclusion}

This specification defines the core rules required
to generate and verify deterministic decision evidence.

\end{document}