\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\title{DIP-003: Append-Only Decision Ledgers\\
Verifiable Evidence Registries}

\author{Pavan Dev Singh Charak}

\date{2026}

\begin{document}

\maketitle

\begin{abstract}

This document describes the Decision Ledger used in the
Decision Integrity Protocol (DIP). The ledger is an append-only
registry that stores signed decision payloads and ensures
integrity of historical decision evidence.

\end{abstract}

\section{Introduction}

Decision systems require reliable methods to store
historical decision records. Traditional logging systems
are often mutable and difficult to verify.

The Decision Ledger provides an append-only evidence
registry for storing signed decision payloads.

\section{Ledger Model}

The Decision Ledger records decision payloads in a
chronologically ordered append-only structure.

Each entry contains:

\begin{itemize}
\item Canonical decision payload
\item Cryptographic signature
\item Ledger timestamp
\end{itemize}

\section{Append-Only Guarantees}

The ledger enforces append-only semantics:

\begin{itemize}
\item Existing records MUST NOT be modified
\item New records MAY only be appended
\item Ledger history MUST remain intact
\end{itemize}

\section{Ledger Entry Example}

\begin{verbatim}
{
  "ledger_index": 152,
  "decision_id": "abc123",
  "timestamp": "2026-03-06T10:30:00Z",
  "payload_hash": "f8c9a...",
  "signature": "ed25519..."
}
\end{verbatim}

\section{Ledger Integrity}

Ledger integrity may be ensured through:

\begin{itemize}
\item cryptographic hashing
\item chained ledger entries
\item independent verification
\end{itemize}

\section{Auditability}

Because the ledger is append-only, auditors can reconstruct
the complete history of decision records and validate
decision integrity.

\section{Conclusion}

The Decision Ledger provides the foundational evidence
registry that enables verifiable decision systems.

\end{document}